Advanced Normalization Tools (ANTS)

### 2009-06-24 Documentation/Tex/ants.tex
updates to documentation and output of programs ....
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@44 fd2da0a4-d968-48df-9cb5-d8c496287cbf

Thus, ANTS enables many operating points from both the transformation and appearance domains such that users may make choices appropriate for their problems. The ANTS command-line syntax, shown in figure~\ref{fig:cmd}, reflects these operating points and the various components that interact in the optimization.
\begin{figure}
\label{fig:cmd}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax.pdf} 
\itkcaption[ANTS Command Line]{The relationship between the variational 
form that defines the optimization goals and the ANTS command line.}
\vspace{-0.1in}
\end{figure}
ANTS may be used to navigate the transformation and similarity metric space. 
\subsection{Use Your Header!} ANTS uses the direction/orientation, spacing
and origin in its definitions of the mapping between images.  That is,
ANTS uses the nifti standard defintions of an image space.  So, the
image orientation (direction), origin and spacing are important!  You
may ``check'' for consistency between two image header definitions by
using: ImageMath 3 Image2repaired.nii.gz CompareHeadersAndImages
Image1.nii.gz Image2.nii.gz \newline This will compare Image2 to
Image1 and ``fix'' the header of Image2, writing out to
Image2repaired.

\subsection{ANTS Transformation Models}
The ANTS toolkit provides a hierarchy of transformations with
adjustable levels of complexity, regularization, degrees of freedom
and behavior as optimizers.  The simplest transformation model is the
rigid and/or affine transform.  The most complex -- and most flexible
-- is a symmetric diffeomorphic transformation based on optimizing and
integrating a time-varying velocity field.  Computation time also
increases with transformation model complexity, but most normalization
needs may be met with under an hour of computation.  
We provide an overview of some of the ANTS models below and 
try to communicate intuition on what one gains/loses with each choice.  
An overview of available models and similarity terms is in Table~\ref{table:chart}.
\begin{table*}
  \centering
    \begin{tabular}{c c c c}
    {\bf Category} & {\bf Transformation, $\phi$} & {\bf Similarity Measures} & {\bf Brief Description} \\
    \toprule     
    \multirow{2}*{\bf Linear}
           & Rigid$^\dagger$ & MI, MSQ & Rigid registration. \\
       {} & Affine$^\dagger$ & MI, MSQ & Affine registration. \\
       \cmidrule(l){2-4}
    \multirow{2}*{\bf Elastic}
           & Deformable & CC, PR, MI, MSQ, PSE & Demons-like algorithm. \\
       {} & DMFFD & CC, PR, MI, MSQ, PSE  & FFD variant. \\
       \cmidrule(l){2-4}
    \multirow{2}*{\bf Diffeo.}
           & Exponential$^\dagger$ &  CC,PR, MI, MSQ, PSE  & $~\min~\velocity(\x)$ \\
       {} & Greedy SyN$^\dagger$ &  CC, PR, MI, MSQ, PSE  & ~~locally in time~$\min~\velocity(\x,t)$\\
       {} & Geodesic SyN$^\dagger$ &  CC, PR, MI, MSQ, PSE  &  $~\min~\velocity(\x,t)$ over all time \\
    \bottomrule
    \end{tabular}
  \caption{Transformations and a subset of the similarity metrics available in ANTS.  Similarity metric acronyms:  CC = fast cross correlation, PR = pure cross correlation (the preferred metric), MSQ = mean squared difference, MI = mutual information, PSE = point-set expectation.  ANTS also provides the inverse of those transformations denoted by the `$\dagger$' symbol.  The brief descriptions of the diffeomorphic algorithms contrast the way in which the velocity field is optimized and used to parameterize $\phi$, the mapping.  All 
ANTS Diff algorithms generate $\phi(\x,t)$ over $t \in [ 0, 1]$ through gradient descent.}
  \label{table:chart}
\end{table*}    
\vspace{0in}

@@

\noindent{\bf Diffeomorphic Transformations.}  The elastic 
mapping space -- as indicated above -- may prove inadequate 
for some large deformation mapping scenarios.  Figures~\ref{fig:antshier} 
illustrate the changing performance one may get in switching from affine 
to elastic to diffeomorphic normalization.  
\begin{figure}
\label{fig:antssyn}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSsyn.pdf} 
\itkcaption[ANTS Symmetric Normalization]{This example shows the benefit of the symmetric normalization model -- invertibility, symmetry, highly deformable and accurate registration.  The transformation model used here is highlighted in red.}
\vspace{-0.1in}
\end{figure}
Figure~\ref{fig:antssyn} shows 
how one may use ANTS to achieve a state-of-the-art diffeomorphic mapping. 
The ANTS diffeomorphic model chosen for this example -- symmetric normalization -- 
is invariant to the order of the input images (although the affine mapping is not).  
An additional advantage of the diffeomorphic model over the elastic model is 
that both forward and inverse transformations are computed thus allowing one 
to deform fixed to moving and also moving to fixed.   The transformation model 
chosen here -- \verb SyN[0.25] -- may be replaced with other 
diffeomorphic transformation models.  The most general is global-in-time SyN via \verb SyN[0.25,2,0.01]   , 
where the time step for integration is 0.01 (lower is more accurate).  A fast approximate 
solution may be gained through 
exponential mapping via \verb Exp[0.25,10] , where 10 integrations points are used. 
Update schemes for diffeomorphic optimization are similar to the elastic case 
and are described elsewhere \cite{Avants2008b,ANTStheory}.

@@

\subsection{Optimal Template Construction with ANTS}
A useful script for setting up optimal template construction
 is available in ANTS/Scripts.  Two versions of available : serial and parallel.  
Data and a script for building an average face template is available 
the ANTS/Examples/Data directory.  Two key points about optimal templates:
1. The outcome stabilizes at around ten images, for most populations.
That is, if we randomly select images from individuals within the same 
demographic distribution, we will end up with very similar average images.  
2.  Optimality, for the ANTS-SyN approach, is defined by the minimum shape 
and appearance distance image. 
See the template page for some examples of this and users may 
contact ANTS developers for previously derived template 
images that may be useful.

@@

\section{Image Segmentation and Labeling}
ANTS has tools for both tissue based segmentation and prior-based segmentation that 
leverages spatial priors, usually based on a template mapping.  
\subsection{Basic Segmentation}
A basic MRF segmentation algorithm is available in ANTS ImageMath.  We apply 
the segmentation to the example data in: ANTS/Examples/Data/r16slice.nii.  
\begin{verbatim}
ImageMath 2 r16slice.nii Segment r16slice.nii 4 0 
% output = r16slice_seg.nii, r16slice_prob_0.nii, r16slice_prob_1.nii, r16slice_prob_2.nii .  
\end{verbatim}
The parameter \verb 4 ~indicates 3 tissues plus background are sought.  The \verb 0 ~indicates 
that there are no prior images available to guide the segmentation.  Output is shown 
in figure~\ref{fig:seg} and -- with priors -- in figure~\ref{fig:seg2}.
\begin{figure}
\label{fig:seg}
\includegraphics[width=1\textwidth]{Figures/segmentation.pdf} 
\itkcaption[ANTS Segmentation]{ANTS provides basic 
Markov Random Field regularized Gaussian model tissue segmentation. 
The ANTS program N3BiasFieldCorrection is very valuable as preprocessing 
for this naive approach to segmentation.   The last panel -- far right -- 
shows the thickness derived from the white matter and gray matter 
probabilities where a prior on thickness was used to prevent thickness
overestimation.   Thickness was derived with DiReCT \cite{Das2009}, a software 
tool available in binary form only -- a future release may provide code.  
}
\end{figure}
\subsection{Prior and Template-based Image Segmentation}
The same algorithm may be augmented 
to perform Prior-based segmentation.  
\begin{verbatim}
ImageMath 2 r16sliceprior.nii Segment r16slice.nii 4 0.5 
   csfprior.nii wmprior.nii  gmprior.nii 
% output = r16sliceprior_seg.nii  r16sliceprior_prob_0.nii  
% r16sliceprior_prob_1.nii  r16sliceprior_prob_2.nii .  
\end{verbatim}
The parameter \verb 4 indicates 3 tissues plus background are sought.  The \verb 0.5 indicates 
we weight the priors equally as the data term and use a spatially varying set of Gaussians 
to estimate the segmentation.  In this way, data from the priors may be used to modify 
and guide the segmentation in a locally varying way, accomodating for both inhomogeneity 
and the different imaging signature that different tissues provide.   Prior images should be 
of the same size and dimension as the input data and should have intensities in the range $[0,1]$.
\begin{figure}
\label{fig:seg2}
\includegraphics[width=1\textwidth]{Figures/segmentation2.pdf}
\itkcaption[ANTS Segmentation]{ANTS extends basic Markov Random Field
  regularized Gaussian model tissue segmentation to include priors,
  which allow spatially varying tissue models to determine
  segmentation.  The ANTS program N3BiasFieldCorrection is less
  critical as preprocessing when using this approach to segmentation.
  Here, we see the improvement in segmentation as the locally varying 
  prior models are weighted more heavily.  
  The last panel -- far right -- shows the segmentation derived from
  data driven with an initialization founded purely on prior models 
  based on template mapping.  The prior model brings out the caudate 
and some CSF -- highlighted by arrows -- as their weight increases. 
}
\end{figure}
\subsection{Cortical Thickness}
Two forms of cortical thickness estimation from probability maps are available 
in ANTS: first, the traditional Laplacian cortical thickness estimation and, second, 
the more recently developed Diffeomorphic Registration-based Cortical Thickness 
(DiReCT) \cite{Das2009}.  Both methods estimate thickness of an object 
from probability or binary images that define object boundaries.  This tool 
is mainly of interest in brain mapping and cardiac imaging for morphometric 
studies.  Cortical thickness, for instance, is known to correlate with language 
development and IQ in adolescents.

\subsection{User/Label-Guided Normalization for Hippocampus Mapping}
See the link for details on hippocampus labeling:
\href{http://picsl.upenn.edu/ANTS/hipptutorial.php}{http://picsl.upenn.edu/ANTS/hipptutorial.php}.
\noindent{\bf Expectation-Based Point-Set Registration.}
Here, we apply the expectation-based point set registration method 
for mapping labeled points sets.   ITK-SNAP may be used to label 
images and exported segmentation images may be input to the 
PSE metric below, as labeled data.  The Frown and Smile data is used 
as example.  This data is available in the 
ANTS/Examples/Data/ directory.
\begin{figure}
\label{fig:frown}
\includegraphics[width=0.45\textwidth]{Figures/frowntosmile.png}
\includegraphics[width=0.45\textwidth]{Figures/smiletofrown.png}
\itkcaption[Frown To Smile]{Expected output for the frown to smile 
shows a smooth, though large deformation.  The grids are overlaid on 
the deformed images. }
\end{figure}
\begin{verbatim}
ANTS 2 -o PSEtest  -i 91x70x55x40x30  -r Gauss[3,0.] -t SyN[0.2]
   -m  PSE[Frown.nii,Smile.nii,Frown.nii,Smile.nii,0.75,0.1,11,0,10] 
   -m  MSQ[Frown.nii,Smile.nii,1,0]    --number-of-affine-iterations 0

@@

\subsection{Statistics with ANTS and ``R''}
{\bf R} is an open-source, cutting-edge statistical package 
used by statisticians world-wide.   ANTS is designed to 
interface with {\bf R} via an ImageMath tool that converts 
image sets into an {\bf R} compatible format.  
Three steps are required.  First, convert image data into 
vector and matrix formats through a mask of your statistical 
region of interest.  Second, read that data into {\bf R} and 
run your statistics.  Third, write out the data from {\bf R} and 
convert back to an image.  Here are the steps, in code examples,
\begin{enumerate}
\item Create a vector for your output: 
\begin{verbatim} 
ImageMath 3 testresultvec.nii ConvertImageSetToMatrix 1 mask.nii image1.nii
\end{verbatim}
Here, we assume all images are called image1.nii, image2.nii, ... , imageN.nii so we 
can use a wildcard to pass to ImageMath and 
create a matrix for all your input data: 
\begin{verbatim} 
 ImageMath 3 datamatrix.nii ConvertImageSetToMatrix 1 mask.nii image*.nii 
\end{verbatim}
\item   Run {\bf R} by reading your data above, performing whatever statistics you want and then 
writing out.  
\item   Convert the test result back to an image.  
\begin{verbatim} 
ImageMath 3 testresult.nii ConvertVectorToImage mask.nii testresultvec.nii
\end{verbatim}
\end{enumerate}
Most statistical requirements may be met with this setup. 
\section{Dependencies and Related Software}
\subsection{Dependencies and Compilation}
ANTS depends on the most recent version of the Insight ToolKit (ITK)
-- www.itk.org.   File types that may be read and written are 
all of those available in itk and a few more \href{http://picsl.upenn.edu/ANTS/ioants.php}{http://picsl.upenn.edu/ANTS/ioants.php}.

>>>>

Thus, ANTS enables many operating points from both the transformation and appearance domains such that users may make choices appropriate for their problems. The ANTS command-line syntax, shown in figure~\ref{fig:cmd}, reflects these operating points and the various components that interact in the optimization.
\begin{figure}
\label{fig:cmd}
\includegraphics[width=0.9\textwidth]{Figures/ANTSSyntax.pdf} 
\itkcaption[ANTS Command Line]{The relationship between the variational 
form that defines the optimization goals and the ANTS command line.  The 
SyN option implements the method in \cite{Avants2008b} and evaluated in 
\cite{Klein2009}.}
\vspace{-0.1in}
\end{figure}
ANTS may be used to navigate the transformation and similarity metric space. 
\subsection{Use Your Header!} ANTS uses the direction/orientation, spacing
and origin in its definitions of the mapping between images.  That is,
ANTS uses the nifti standard defintions of an image space.  So, the
image orientation (direction), origin and spacing are important!  You
may ``check'' for consistency between two image header definitions by
using: ImageMath 3 Image2repaired.nii.gz CompareHeadersAndImages
Image1.nii.gz Image2.nii.gz \newline This will compare Image2 to
Image1 and ``fix'' the header of Image2, writing out to
Image2repaired.

\subsection{ANTS Transformation Models}
The ANTS toolkit provides a hierarchy of transformations with
adjustable levels of complexity, regularization, degrees of freedom
and behavior as optimizers.  The simplest transformation model is the
rigid and/or affine transform.  The most complex -- and most flexible
-- is a symmetric diffeomorphic transformation based on optimizing and
integrating a time-varying velocity field.  Computation time also
increases with transformation model complexity, but most normalization
needs may be met with under an hour of computation.  
We provide an overview of some of the ANTS models below and 
try to communicate intuition on what one gains/loses with each choice.  
An overview of available models and similarity terms is in Table~\ref{table:chart}.
\begin{table*}
  \centering
    \begin{tabular}{c c c c}
    {\bf Category} & {\bf Transformation, $\phi$} & {\bf Similarity Measures} & {\bf Brief Description} \\
    \toprule     
    \multirow{2}*{\bf Linear}
           & Rigid$^\dagger$ & MI, MSQ & Rigid registration. \\
       {} & Affine$^\dagger$ & MI, MSQ & Affine registration. \\
       \cmidrule(l){2-4}
    \multirow{2}*{\bf Elastic}
           & Deformable & CC, PR, MI, MSQ, PSE & Demons-like algorithm. \\
       {} & DMFFD & CC, PR, MI, MSQ, PSE  & FFD variant. \\
       \cmidrule(l){2-4}
    \multirow{2}*{\bf Diffeo.}
           & Exponential$^\dagger$ &  CC,PR, MI, MSQ, PSE  & $~\min~\velocity(\x)$ \\
       {} & Greedy SyN$^\dagger$ &  CC, PR, MI, MSQ, PSE  & ~~locally in time~$\min~\velocity(\x,t)$\\
       {} & Geodesic SyN$^\dagger$ &  CC, PR, MI, MSQ, PSE  &  $~\min~\velocity(\x,t)$ over all time \\
    \bottomrule
    \end{tabular}
  \caption{Transformations and a subset of the similarity metrics available in ANTS.  Similarity metric acronyms:  CC = fast cross correlation, PR = pure cross correlation (the preferred metric), MSQ = mean squared difference, MI = mutual information, PSE = point-set expectation \cite{Pluta2008}.  ANTS also provides the inverse of those transformations denoted by the `$\dagger$' symbol.  The brief descriptions of the diffeomorphic algorithms contrast the way in which the velocity field is optimized and used to parameterize $\phi$, the mapping.  All 
ANTS Diff algorithms generate $\phi(\x,t)$ over $t \in [ 0, 1]$ through gradient descent.}
  \label{table:chart}
\end{table*}    
\vspace{0in}

@@

\noindent{\bf Diffeomorphic Transformations.}  The elastic 
mapping space -- as indicated above -- may prove inadequate 
for some large deformation mapping scenarios.  Figures~\ref{fig:antshier} 
illustrate the changing performance one may get in switching from affine 
to elastic to diffeomorphic normalization.  
\begin{figure}
\label{fig:antssyn}
\includegraphics[width=0.9\textwidth]{./Figures/ANTSsyn.pdf} 
\itkcaption[ANTS Symmetric Normalization]{This example shows the benefit of the symmetric normalization model -- invertibility, symmetry, highly deformable and accurate registration.  The transformation model used here is highlighted in red.}
\vspace{-0.1in}
\end{figure}
Figure~\ref{fig:antssyn} shows 
how one may use ANTS to achieve a state-of-the-art diffeomorphic mapping. 
The ANTS diffeomorphic model chosen for this example -- symmetric normalization \cite{Avants2008b,Kim2008} -- 
is invariant to the order of the input images (although the affine mapping is not).  
An additional advantage of the diffeomorphic model over the elastic model is 
that both forward and inverse transformations are computed thus allowing one 
to deform fixed to moving and also moving to fixed.   The transformation model 
chosen here -- \verb SyN[0.25] -- may be replaced with other 
diffeomorphic transformation models.  The most general is global-in-time SyN via \verb SyN[0.25,2,0.01]   , 
where the time step for integration is 0.01 (lower is more accurate).  A fast approximate 
solution may be gained through 
exponential mapping via \verb Exp[0.25,10] , where 10 integrations points are used. 
Update schemes for diffeomorphic optimization are similar to the elastic case 
and are described elsewhere \cite{Avants2008b,ANTStheory}.

@@

\subsection{Optimal Template Construction with ANTS}
A useful script for setting up optimal template construction
 is available in ANTS/Scripts.  This method is described in \cite{Avants2004,Avants2006d,Kim2008,Yushkevich2009}, 
where more recent publications are more descriptive and incorporate 
more recent features and optimization strategies. 
Two versions of this algorithm are available : serial and parallel.  
Data and a script for building an average face template is available 
the ANTS/Examples/Data directory.  Two key points about optimal templates:
1. The outcome stabilizes at around ten images, for most populations.
That is, if we randomly select images from individuals within the same 
demographic distribution, we will end up with very similar average images.  
2.  Optimality, for the ANTS-SyN approach, is defined by the minimum shape 
and appearance distance image. 
See the template page for some examples of this and users may 
contact ANTS developers for previously derived template 
images that may be useful.

@@

\section{Image Segmentation and Labeling}
ANTS has tools for both tissue based segmentation and prior-based segmentation that 
leverages spatial priors, usually based on a template mapping.  
\subsection{Basic Segmentation}
A basic MRF segmentation algorithm is available in ANTS ImageMath.  We apply 
the segmentation to the example data in: ANTS/Examples/Data/r16slice.nii.  
ImageMath 2 r16slice.nii Segment r16slice.nii 4 0 
% output = r16slice_seg.nii, r16slice_prob_0.nii, r16slice_prob_1.nii, r16slice_prob_2.nii .  
The parameter \verb 4 ~indicates 3 tissues plus background are sought.  The \verb 0 ~indicates 
that there are no prior images available to guide the segmentation.  Output is shown 
in figure~\ref{fig:seg} and -- with priors -- in figure~\ref{fig:seg2}.
\label{fig:seg}
\includegraphics[width=1\textwidth]{Figures/segmentation.pdf} 
\itkcaption[ANTS Segmentation]{ANTS provides basic 
Markov Random Field regularized Gaussian model tissue segmentation. 
The ANTS program N3BiasFieldCorrection is very valuable as preprocessing 
for this naive approach to segmentation.   The last panel -- far right -- 
shows the thickness derived from the white matter and gray matter 
probabilities where a prior on thickness was used to prevent thickness
overestimation.   Thickness was derived with DiReCT \cite{Das2009}, a software 
tool available in binary form only -- a future release may provide code.  
\subsection{Prior and Template-based Image Segmentation}
The same algorithm may be augmented 
to perform Prior-based segmentation.  
ImageMath 2 r16sliceprior.nii Segment r16slice.nii 4 0.5 
   csfprior.nii wmprior.nii  gmprior.nii 
% output = r16sliceprior_seg.nii  r16sliceprior_prob_0.nii  
% r16sliceprior_prob_1.nii  r16sliceprior_prob_2.nii .  
The parameter \verb 4 indicates 3 tissues plus background are sought.  The \verb 0.5 indicates 
we weight the priors equally as the data term and use a spatially varying set of Gaussians 
to estimate the segmentation.  In this way, data from the priors may be used to modify 
and guide the segmentation in a locally varying way, accomodating for both inhomogeneity 
and the different imaging signature that different tissues provide.   Prior images should be 
of the same size and dimension as the input data and should have intensities in the range $[0,1]$.
\label{fig:seg2}
\includegraphics[width=1\textwidth]{Figures/segmentation2.pdf}
\itkcaption[ANTS Segmentation]{ANTS extends basic Markov Random Field
  regularized Gaussian model tissue segmentation to include priors,
  which allow spatially varying tissue models to determine
  segmentation.  The ANTS program N3BiasFieldCorrection is less
  critical as preprocessing when using this approach to segmentation.
  Here, we see the improvement in segmentation as the locally varying 
  prior models are weighted more heavily.  
  The last panel -- far right -- shows the segmentation derived from
  data driven with an initialization founded purely on prior models 
  based on template mapping.  The prior model brings out the caudate 
and some CSF -- highlighted by arrows -- as their weight increases. 
\subsection{Cortical Thickness}
Two forms of cortical thickness estimation from probability maps are available 
in ANTS: first, the traditional Laplacian cortical thickness estimation and, second, 
the more recently developed Diffeomorphic Registration-based Cortical Thickness 
(DiReCT) \cite{Das2009}.  Both methods estimate thickness of an object 
from probability or binary images that define object boundaries.  This tool 
is mainly of interest in brain mapping and cardiac imaging for morphometric 
studies.  Cortical thickness, for instance, is known to correlate with language 
development and IQ in adolescents.  The laplacian approach to 
cortical thickness may be used via the program LaplacianThickness, 
which implements the method described in ``Three-dimensional mapping of cortical thickness using Laplace's Equation'' by 
Jones, et al \cite{}.  Super-sampling 
and controlling segmentation accuracy for input to this program is up 
to the user.  Otherwise, the Laplacian method may grossly overestimate 
thickness in closed sulci.

\subsection{User/Label-Guided Normalization for Hippocampus Mapping}
See the link for details on hippocampus labeling:
\href{http://picsl.upenn.edu/ANTS/hipptutorial.php}{http://picsl.upenn.edu/ANTS/hipptutorial.php}.
\noindent{\bf Expectation-Based Point-Set Registration.}
Here, we apply the expectation-based point set registration method 
for mapping labeled points sets, as described in \cite{Pluta2008}.   ITK-SNAP may be used to label 
images and exported segmentation images may be input to the 
PSE metric below, as labeled data.  The Frown and Smile data is used 
as example.  This data is available in the 
ANTS/Examples/Data/ directory.
\begin{figure}
\label{fig:frown}
\includegraphics[width=0.45\textwidth]{Figures/frowntosmile.png}
\includegraphics[width=0.45\textwidth]{Figures/smiletofrown.png}
\itkcaption[Frown To Smile]{Expected output for the frown to smile 
shows a smooth, though large deformation.  The grids are overlaid on 
the deformed images. }
\end{figure}
\begin{verbatim}
ANTS 2 -o PSEtest  -i 91x70x55x40x30  -r Gauss[3,0.] -t SyN[0.2]
   -m  PSE[Frown.nii,Smile.nii,Frown.nii,Smile.nii,0.75,0.1,11,0,10] 
   -m  MSQ[Frown.nii,Smile.nii,1,0]    --number-of-affine-iterations 0

@@

\subsection{Statistics with ANTS and {\bf R}}
{\bf R} is an open-source, cutting-edge statistical package 
used by statisticians world-wide.   ANTS is designed to 
interface with {\bf R} via an ImageMath tool that converts 
image sets into an {\bf R} compatible format.  
Three steps are required.  First, convert image data into 
vector and matrix formats through a mask of your statistical 
region of interest.  Second, read that data into {\bf R} and 
run your statistics.  Third, write out the data from {\bf R} and 
convert back to an image.  Here are the steps, in code examples,
\begin{enumerate}
\item Create a vector for your output: 
\begin{verbatim} 
ImageMath 3 testresultvec.nii ConvertImageSetToMatrix 1 mask.nii image1.nii
\end{verbatim}
Here, we assume all images are called image1.nii, image2.nii, ... , imageN.nii so we 
can use a wildcard to pass to ImageMath and 
create a matrix for all your input data: 
\begin{verbatim} 
 ImageMath 3 datamatrix.nii ConvertImageSetToMatrix 1 mask.nii image*.nii 
\end{verbatim}
\item   Run {\bf R} by reading your data above, performing whatever statistics you want and then 
writing out.  
\item   Convert the test result back to an image.  
\begin{verbatim} 
ImageMath 3 testresult.nii ConvertVectorToImage mask.nii testresultvec.nii
\end{verbatim}
\end{enumerate}
Most statistical requirements may be met with this setup. 
\section{Dependencies and Related Software}
\subsection{Dependencies and Compilation}
ANTS depends on the most recent version of the Insight ToolKit (ITK)
-- www.itk.org.   File types that may be read and written are 
all of those available in itk and a few more \href{http://picsl.upenn.edu/ANTS/ioants.php}{http://picsl.upenn.edu/ANTS/ioants.php}.
